Policy optimization for dynamic spatiotemporal systems \hspace{1em} \insertframenumber

### 2014-08-04 pot/LaTeX/jsm2014.tex
typo correction

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item \alert{Conditioned} on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is Gaussian
    \item Can use standard geostatistical methods to fit this model.
    \item Predictions can be made through kriging.
    \item \alert{Marginalizing} over $z_t$ and $\sigma^2_t$ (via MCMC),
    \begin{align*}
      Y_t(\bs) \sim \text{skew-t}(\mu, \alpha, \text{df}=2a)
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Long-range dependence}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item The $\chi$ coefficient is a measure of extremal spatial correlation
    \begin{align*}
      \chi(\bh) = \Pr(Y_t(\bs) > c \mid Y_t(\bs + \bh) > c)
    \end{align*}
    \item This value shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$.
    \item Deal with this through a daily random partition.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 9 different analysis methods incorporating
    \begin{itemize}
      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
      \item $K=1$ partition vs $K=5$ partitions
      \item No thresholding vs thresholded
      \begin{itemize}
        \item Thresholded data at $T=0.90$ sample quantile
      \end{itemize}
    \end{itemize}
    \item All methods use a \Matern or exponential covariance ($\nu = 0.5$)
    \item Compare quantile and Brier scores using 5-fold cross validation
    \item Mean function modeled using a first-order spatial trend
  \end{itemize}
\end{frame}

\begin{frame}{Cross-validation results}
  \begin{table}[htbp]
    \small
    \centering
    \begin{tabular}{l|c|r|rrrrr}
          \multicolumn{3}{c}{\ } & \multicolumn{5}{c}{Quantile}\\
           \hline
  Marginal & $K$ & $T$  & 0.900 & 0.950 & 0.990 & 0.995 & 0.999\\
  \hline
Gaussian & 1 & 0 & 16.38 & 15.76 & 14.52 & 14.08 & 13.22\\
$t$ & 1 & 0 & 16.15 & 15.51 & 14.00 & 13.43 & 12.32\\
$t$ & 5 & 0 & 13.61 & 12.66 & 10.96 & 10.40 & 9.34\\
skew $t$ & 1 & 0 & 9.24  & 7.27 & 4.13  & 3.27  & 1.96\\
skew $t$ & 5 & 0 & 15.81 & 14.46 & 11.57 & 10.57 & 8.60\\
$t$ & 1 & 0.9 & 5.52  & 3.58  & 1.77  & 1.47  & 1.10\\
$t$ & 5 & 0.9 & 5.98  & 4.27  & 2.41  & 2.03  & 1.49\\
 skew $t$ & 1 & 0.9 & {\bf 4.91}  & {\bf 3.16} & {\bf 1.45}  & {\bf 1.16}  & {\bf 0.82}\\
skew $t$ & 3 & 0.9 & 5.58 & 3.78 & 1.93& 1.58& 1.11\\
\hline
    \end{tabular}
  \end{table}
\end{frame}

@@

\begin{frame}{Simulation results}
  Still rerunning a few of the MCMC runs because quantile scores for 4. and 5. are
  unusually high for method 5
\end{frame}

@@

\begin{frame}{References}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Cooley, D., Cisewski, J., Erhardt, R. J., Mannshardt, E., Omolo, B. O. and Sun, Y. (2012) A survey of spatial extremes: Measuring spatial dependence and modeling spatial effects. {\it REVSTAT}, {\bf 10}, 135--165.
    \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461
    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21} 33--47.
  \end{itemize}
\end{frame}

>>>>

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item \alert{Conditioned} on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is Gaussian
    \item Can use standard geostatistical methods to fit this model.
    \item Predictions can be made through kriging.
    \item \alert{Marginalizing} over $z_t$ and $\sigma^2_t$ (via MCMC),
    \begin{align*}
      Y_t(\bs) \sim \text{skew-t}(\mu, \Sigma^*, \alpha, \text{df}=2a)
    \end{align*}
    where
    \begin{itemize}
    	\item $\mu$ is the location
	\item $a$, $b$ are the IG parameters for $\sigma^2_t$
	\item $\Sigma^* = \frac{ b }{ a } \Sigma$ is a scale matrix, and $\Sigma$ is a \Matern covariance matrix
	\item $\alpha \in \calR$ controls the skewness
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Long-range dependence}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item The $\chi$ coefficient is a measure of extremal spatial correlation
    \begin{align*}
      \chi(\bh) = \Pr(Y_t(\bs) > c \mid Y_t(\bs + \bh) > c)
    \end{align*}
    \item This value shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$ (Padoan, 2011)
    \item Deal with this through a daily random partition.
  \end{itemize}
\end{frame}

\begin{frame}{Simulated $\chi$ plots}
  \centering
  \includegraphics[width=1\linewidth]{./plots/chi-plots.pdf}\\[-0.25in]
  \includegraphics[width=0.2\linewidth]{./plots/chi-legend.pdf}
\end{frame}

@@

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 9 different analysis methods incorporating
    \begin{itemize}
      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
      \item $K=1$ partition vs $K=5$ partitions
      \item No thresholding vs thresholded
      \begin{itemize}
        \item Thresholded data at $T=0.90$ sample quantile
      \end{itemize}
    \end{itemize}
    \item All methods use a \Matern or exponential covariance ($\nu = 0.5$)
    \item Compare quantile and Brier scores using 5-fold cross validation (Gneiting and Raftery, 2007)
    \item Mean function modeled using a first-order spatial trend
  \end{itemize}
\end{frame}

\begin{frame}{Quantile score}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item The quantile score for the $\tau$th quantile is
    \begin{align*}
      2 \{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q} - y)
    \end{align*}
    where:
    \begin{itemize}
      \item $y$ is a test set value
      \item $\widehat{q}(\tau)$ is the estimated $\tau$th quantile
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Brier score}
  \begin{itemize} \setlength{\itemsep}{0.5em}
	\item Brier score for predicting exceedance of threshold $c$
	\begin{align*}
	  [e(c) - P(c)]^2
	\end{align*}
	where 
	\begin{itemize}
		\item $y$ is a test set value
		\item $e(c) = I[y > c]$
		\item $P(c)$ is the predicted probability of exceeding $c$
	\end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Five-fold cross-validation results}
  \begin{table}[htbp]
    \small
    \centering
    \begin{tabular}{l|c|r|rrrrr}
          \multicolumn{3}{c}{\ } & \multicolumn{5}{c}{Quantile}\\
           \hline
  Marginal & $K$ & $T$  & 0.900 & 0.950 & 0.990 & 0.995 & 0.999\\
  \hline
Gaussian & 1 & 0 & 16.38 & 15.76 & 14.52 & 14.08 & 13.22\\
$t$ & 1 & 0 & 16.15 & 15.51 & 14.00 & 13.43 & 12.32\\
$t$ & 5 & 0 & 13.61 & 12.66 & 10.96 & 10.40 & 9.34\\
skew $t$ & 1 & 0 & 9.24  & 7.27 & 4.13  & 3.27  & 1.96\\
skew $t$ & 5 & 0 & 15.81 & 14.46 & 11.57 & 10.57 & 8.60\\
$t$ & 1 & 0.9 & 5.52  & 3.58  & 1.77  & 1.47  & 1.10\\
$t$ & 5 & 0.9 & 5.98  & 4.27  & 2.41  & 2.03  & 1.49\\
 skew $t$ & 1 & 0.9 & {\bf 4.91}  & {\bf 3.16} & {\bf 1.45}  & {\bf 1.16}  & {\bf 0.82}\\
skew $t$ & 3 & 0.9 & 5.58 & 3.78 & 1.93& 1.58& 1.11\\
\hline
    \end{tabular}
  \end{table}
  \begin{itemize}
  	\item Brier score results are similar.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Simulation study results}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Results are similar to the results from the data analysis
    \item Biggest gains come from thresholding.
    \item Using skew models give additional gain, but small relative to gain for thresholding.
  \end{itemize}
\end{frame}

@@

\begin{frame}{References}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461.
    \item Padoan, S. A. (2011) Multivariate extreme models based on underlying skew-$t$ and skew-normal distributions. {\it Journal of Multivariate Analysis}, {\bf 102}, 977--991.
    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21}, 33--47.
  \end{itemize}
\end{frame}
